### 2015-03-13 article/article.tex
commit

%------------------------------------------------
\subsection{Nuclear receptors in the human genome}
The nuclear receptor variation has been widely studied not only in the mouse, but also in the human genome. The research hereby refers to two published genome-wide association studies in this field of interest, namely \textit{Kora}\footnote{\url{http://epi.helmholtz-muenchen.de/kora-gen}, \today}  - a GWAS study initiated at the Helmholtz Zentrum in Munich involving approximately 18,000 adults from Southern Germany - and \textit{TwinsUK}\footnote{\url{http://www.twinsuk.ac.uk/}, \today}, being the biggest adult twin genetic registry in the UK.  
~~~~~~~\\
~~~~~~~\\
Using the information provided by these studies regarding the correlations between different SNPs and metabolites to metabolic pathways, this research aimed to identify the similarities between the nuclear receptors and their associated phenotypes in human and in the mouse, respectively. Due to missing or incomplete annotations for the nuclear receptors in the mouse, comparisons could only be made between the phenotypes extracted from the MGI and MPD databases (see tables \textit{mgi\_phenotypes} and \textit{mpg\_phenotypes} from the internal database~\ref{fig:database}). Table \ref{tab:compHuMo} shows a list of nuclear receptors found in both the mouse and human and which present a similar phenotype or share the same biological function.
~~~~~~~\\
~~~~~~~\\
%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%
%which are the joined phenotypes, what's the consequence? e.g. what does it mean that mouse + human share steroid hormone? 
%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%
~~~~~~~\\
\begin{table}[H]
\begin{tabulary}{\linewidth}{ LLL }
\hline
	\textbf{Nuclear Receptor Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
	\\  
	\hline \hline 
	Nr2e1  & corticosterone, \textit{steroid hormone} & lathosterol, \textit{steroid hormone}\\
	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
	Thrb  & \textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
	Vdr & \textit{osteoporosis} & \textit{osteoporosis}\\
	Ar & \textit{spinal and bulbar muscular atrophy} & \textit{spinal and bulbar muscular atrophy}\\
	\hline
\end{tabulary}
\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
~~~~~~~\\
Nuclear receptors in human and in the mouse which share a similar phenotype or biological function.}} 
\label{tab:compHuMo}
\end{table}

@@

Over the last year, the number of studies and publications involving the nuclear receptor variation has increased and scientist have been focusing on exploring the effects of nuclear receptors in different organisms, on various levels: in the cells, tissues, and even full-body impact. The goal hereby is to link their metabolic activity to possible human diseases and facilitate the drug development processes. In this regard, extensive experiments and studies have been performed on humans, as well as on laboratory mice, which are known to present a striking genomic resemblance to humans (up to 95\%) and and time and money-wise more advantageous.   As a result, it has been proved that nuclear receptors have a great influence over the extreme phenotypes showcased in the mouse, as well as in human. 
~~~~~~~\\
~~~~~~~\\
In spite of the ascending progress in this field of research, there annotation data provided in the mouse public databases - \textit{Mouse Genome Informatics} and \textit{Mouse Phenome Database} on nuclear receptors is still rather limited. However, the two databases provide complementary information: the phenotypes annotated in \textit{MGI} describe roughly the biological systems in which they occur rather than the description of differences between non-variants, as of \textit{MPD}. On the other hand, the UCSC database for the human genome is far more detailed, comprising also information regarding the metabolic pathways associated with various phenotypes. 
~~~~~~~\\
~~~~~~~\\
This research presents the most significant phenotypes found in the mouse, based on the 49 nuclear receptor variation, as well as a mapping of these phenotypes on the human phenotypic expression, providing insight into possible human disease conditions. Nevertheless, the mouse public databases are far less comprehensive than the UCSC, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%what are the similarites, differences? How does the mouse study involve human phenotypes?
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%------------------------------------------------
\subsection{Nuclear receptors in the human genome}
The nuclear receptor variation has been widely studied not only in the mouse, but also in the human genome. The research hereby refers to two published genome-wide association studies in this field of interest, namely \textit{Kora}\footnote{\url{http://epi.helmholtz-muenchen.de/kora-gen}, \today}  - a GWAS study initiated at the Helmholtz Zentrum in Munich involving approximately 18,000 adults from Southern Germany - and \textit{TwinsUK}\footnote{\url{http://www.twinsuk.ac.uk/}, \today}, being the biggest adult twin genetic registry in the UK.  
~~~~~~~\\
~~~~~~~\\
Using the information provided by these studies regarding the correlations between different SNPs and metabolites to metabolic pathways, this research aimed to identify the similarities between the nuclear receptors and their associated phenotypes in human and in the mouse, respectively. Due to missing or incomplete annotations for the nuclear receptors in the mouse, comparisons could only be made between the phenotypes extracted from the MGI and MPD databases (see tables \textit{mgi\_phenotypes} and \textit{mpg\_phenotypes} from the internal database~\ref{fig:database}). Table \ref{tab:compHuMo} shows a list of nuclear receptors found in both the mouse and human and which present a similar phenotype or share the same biological function.
As seen there are only a few phenotypes shared between these two species. The reason therefore is that the phenotypes for the nuclear receptors in both, human and mouse do not only have influence on one metabolic pathway. For each nuclear receptor there were roughly about 8 to 10 phenotype annotations available, but only the ones which share the same function were written into the table. Nevertheless as an example the gene \textit{Nr2e1} without limitation regulates the expression for the same kind of molecule, a steroid hormone. This leads to the assumption that the nuclear receptors have somehow a similar influence on the individual in each species, albeit there are only few annotated by this time.
~~~~~~~\\
~~~~~~~\\
~~~~~~~\\
\begin{table}[H]
\begin{tabulary}{\linewidth}{ LLL }
\hline
	\textbf{Nuclear Receptor Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
	\\  
	\hline \hline 
	Nr2e1  & corticosterone, \textit{steroid hormone} & lathosterol, \textit{steroid hormone}\\
	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
	Thrb  & \textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
	Vdr & \textit{osteoporosis} & \textit{osteoporosis}\\
	Ar & \textit{spinal and bulbar muscular atrophy} & \textit{spinal and bulbar muscular atrophy}\\
	\hline
\end{tabulary}
\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
~~~~~~~\\
Nuclear receptors in human and in the mouse which share a similar phenotype or biological function.}} 
\label{tab:compHuMo}
\end{table}

@@

Over the last year, the number of studies and publications involving the nuclear receptor variation has increased and scientist have been focusing on exploring the effects of nuclear receptors in different organisms, on various levels: in the cells, tissues, and even full-body impact. The goal hereby is to link their metabolic activity to possible human diseases and facilitate the drug development processes. In this regard, extensive experiments and studies have been performed on humans, as well as on laboratory mice, which are known to present a striking genomic resemblance to humans (up to 95\%) and and time and money-wise more advantageous. As a result, it has been proved that nuclear receptors have a great influence over the extreme phenotypes showcased in the mouse, as well as in human. 
~~~~~~~\\
~~~~~~~\\
In spite of the ascending progress in this field of research, there annotation data provided in the mouse public databases - \textit{Mouse Genome Informatics} and \textit{Mouse Phenome Database} on nuclear receptors is still rather limited. However, the two databases provide complementary information: the phenotypes annotated in \textit{MGI} describe roughly the biological systems in which they occur rather than the description of differences between non-variants, as of \textit{MPD}. On the other hand, the UCSC database for the human genome is far more detailed, comprising also information regarding the metabolic pathways associated with various phenotypes. 
~~~~~~~\\
~~~~~~~\\
This research presents the most significant phenotypes found in the mouse, based on the 49 nuclear receptor variation, as well as a mapping of these phenotypes on the human phenotypic expression, providing insight into possible human disease conditions. Nevertheless, the mouse public databases are far less comprehensive than the UCSC, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
~~~~~~~\\
~~~~~~~\\
The mapping of phenotype information between mouse and human showed that regardless if the genes are thomolog, their function can differ between these two species, at least with the information we have by this time on the human nuclear receptors. As long as the human nuclear receptors are not enough researched, mapping the nuclear receptor variatons from mouse on the human can only result in an indicator to get information about what protein the gene could code for or in which biological process is it likely to be involved.
